### 2017-11-16 Draft2/Appending.Compare.tex
Thank god

In this section, $\FuncWritesDynamicArray$ and $\FuncWritesGrowthArray$ will denote the write cost functions for dynamic and growth arrays, respectively. My goal is to approximate the ratio $\dfrac{\FuncTimeDynamicArray(n)}{\FuncTimeGrowthArray(n)}$ for large $n$. From my assumption near the beginning of Section \ref{subsec:Appending}, this is equal to $\dfrac{\FuncWritesDynamicArray(n)}{\FuncWritesGrowthArray(n)}$. I derived earlier that
\begin{align*}
\left( \frac{\VarGrowthFactor}{\VarGrowthFactor - 1} \right) n \FluteLeq \FuncWritesDynamicArray(n) &\FluteLess \left( \frac{2\VarGrowthFactor - 1}{\VarGrowthFactor - 1} \right) n
\end{align*}
From Theorem \ref{thm:BothSidesInequality}, I may divide all sides of this inequality by $n$ to receive
\begin{align*}
\frac{\VarGrowthFactor}{\VarGrowthFactor - 1} \FluteLeq \frac {\FuncWritesDynamicArray(n)} n &\FluteLess \frac{2\VarGrowthFactor - 1}{\VarGrowthFactor - 1}
\end{align*}
It was shown earlier that $\FuncWritesGrowthArray(n) \sim n$. From Theorem \ref{thm:MergesOverOps}, $\dfrac {\FuncWritesDynamicArray(n)} n \sim \dfrac{\FuncWritesDynamicArray(n)}{\FuncWritesGrowthArray(n)}$. From Theorem \ref{thm:InterchangeableInInequality}, it follows that
\begin{align*}
\frac{\VarGrowthFactor}{\VarGrowthFactor - 1} \FluteLeq \frac {\FuncWritesDynamicArray(n)} {\FuncWritesGrowthArray(n)} &\FluteLess \frac{2\VarGrowthFactor - 1}{\VarGrowthFactor - 1}
\end{align*}
For $\VarGrowthFactor = 2$, this can be interpreted as ``For large $n$, dynamic arrays make approximately $2$ to $3$ times as many writes as growth arrays. \underline{Thus, for large $n$, they are about $2$ to $3$ times slower than growth arrays.}''

@@

Let $\FuncSpaceDynamicArray(n)$ and $\FuncSpaceGrowthArray(n)$ be the space cost functions for dynamic and growth arrays, respectively. I wish to approximate the ratio $\dfrac{\FuncSpaceDynamicArray(n)}{\FuncSpaceGrowthArray(n)}$ for large $n$. Recall that
\begin{align*}
\FuncSpaceDynamicArray(n) = \left( \frac{\VarGrowthFactor^{\VarUseful + 1} - 1}{\VarGrowthFactor - 1} \right) \VarInitCapacity
\end{align*}
and that $\FuncSpaceGrowthArray(n) \sim \VarCapacitySeq_\VarUseful = \VarGrowthFactor^\VarUseful\VarInitCapacity$. Dividing $\FuncSpaceDynamicArray(n)$ by $\FuncSpaceGrowthArray(n)$, I receive
\begin{align*}
\frac{\FuncSpaceDynamicArray(n)}{\FuncSpaceGrowthArray(n)} &\sim \frac{1}{\VarGrowthFactor^\VarUseful\VarInitCapacity} \cdot \left( \frac{\VarGrowthFactor^{\VarUseful + 1} - 1}{\VarGrowthFactor - 1} \right) \VarInitCapacity &&\text{(Theorem \ref{thm:MergesOverOps})}\\
&= \frac{\VarGrowthFactor - 1 / {\VarGrowthFactor^\VarUseful}}{\VarGrowthFactor - 1}\\
&\sim \frac{\VarGrowthFactor}{\VarGrowthFactor - 1} &&\text{(Theorem \ref{thm:RemovesLowerOrderTerms})}
\end{align*}
For $\VarGrowthFactor = 2$, this can be interpreted as ``\underline{For large $n$, dynamic arrays use approximately twice as much space as growth arrays.}''

>>>>

In this section, $\FuncWritesDynamicArray$ and $\FuncWritesGrowthArray$ will denote the write cost functions for dynamic and growth arrays, respectively. My goal is to approximate the ratio $\frac{\FuncTimeDynamicArray(n)}{\FuncTimeGrowthArray(n)}$ for large $n$. From my assumption near the beginning of Section \ref{subsec:Appending}, this is equal to $\frac{\FuncWritesDynamicArray(n)}{\FuncWritesGrowthArray(n)}$. I derived earlier that
\begin{align*}
\left( \frac{\VarGrowthFactor}{\VarGrowthFactor - 1} \right) n \FluteLeq \FuncWritesDynamicArray(n) &\FluteLess \left( \frac{2\VarGrowthFactor - 1}{\VarGrowthFactor - 1} \right) n
\end{align*}
From Theorem \ref{thm:BothSidesInequality}, I may divide all sides of this inequality by $n$ to receive
\begin{align*}
\frac{\VarGrowthFactor}{\VarGrowthFactor - 1} \FluteLeq \frac {\FuncWritesDynamicArray(n)} n &\FluteLess \frac{2\VarGrowthFactor - 1}{\VarGrowthFactor - 1}
\end{align*}
It was shown earlier that $\FuncWritesGrowthArray(n) \sim n$. From Theorem \ref{thm:MergesOverOps}, $\frac {\FuncWritesDynamicArray(n)} n \sim \frac{\FuncWritesDynamicArray(n)}{\FuncWritesGrowthArray(n)}$. From Theorem \ref{thm:InterchangeableInInequality}, it follows that
\begin{align*}
\frac{\VarGrowthFactor}{\VarGrowthFactor - 1} \FluteLeq \frac {\FuncWritesDynamicArray(n)} {\FuncWritesGrowthArray(n)} &\FluteLess \frac{2\VarGrowthFactor - 1}{\VarGrowthFactor - 1}
\end{align*}
For $\VarGrowthFactor = 2$, this can be interpreted as ``For large $n$, dynamic arrays make approximately $2$ to $3$ times as many writes as growth arrays. \underline{Thus, for large $n$, they are about $2$ to $3$ times slower than growth arrays.}''

@@

Let $\FuncSpaceDynamicArray(n)$ and $\FuncSpaceGrowthArray(n)$ be the space cost functions for dynamic and growth arrays, respectively. I wish to approximate the ratio $\frac{\FuncSpaceDynamicArray(n)}{\FuncSpaceGrowthArray(n)}$ for large $n$. Recall that
\begin{align*}
\FuncSpaceDynamicArray(n) = \left( \frac{\VarGrowthFactor^{\VarUseful + 1} - 1}{\VarGrowthFactor - 1} \right) \VarInitCapacity
\end{align*}
and that $\FuncSpaceGrowthArray(n) \sim \VarCapacitySeq_\VarUseful = \VarGrowthFactor^\VarUseful\VarInitCapacity$. Dividing $\FuncSpaceDynamicArray(n)$ by $\FuncSpaceGrowthArray(n)$, I receive
\begin{align*}
\frac{\FuncSpaceDynamicArray(n)}{\FuncSpaceGrowthArray(n)} &\sim \frac{1}{\VarGrowthFactor^\VarUseful\VarInitCapacity} \cdot \left( \frac{\VarGrowthFactor^{\VarUseful + 1} - 1}{\VarGrowthFactor - 1} \right) \VarInitCapacity &&\text{(Theorem \ref{thm:MergesOverOps})}\\
&= \frac{\VarGrowthFactor - 1 / {\VarGrowthFactor^\VarUseful}}{\VarGrowthFactor - 1}\\
&\sim \frac{\VarGrowthFactor}{\VarGrowthFactor - 1} &&\text{(Theorem \ref{thm:RemovesLowerOrderTerms})}
\end{align*}
For $\VarGrowthFactor = 2$, this can be interpreted as ``\underline{For large $n$, dynamic arrays use approximately twice as much space as growth arrays.}''

### 2017-11-16 Draft2/Appending.Dynamic.tex
Thank god

I wish to find the ratio $\dfrac{\FuncTimeDynamicArray(n)}{\FuncTimeGrowthArray(n)}$, where $\FuncTimeDynamicArray(n)$ and $\FuncTimeGrowthArray(n)$ denote the average time needed to append $n$ items for dynamic and growth arrays, respectively. This represents how much faster growth arrays are at appending than dynamic ones. I will approximate this ratio using the following question: Suppose $n$ elements are appended to an empty collection. How many times is an element stored in an array? I will term the answer to this question the \textbf{write cost} of $n$ appends, and denote it $\FuncWrites(n)$.

I will assume that $\FuncTime(n) \propto \FuncWrites(n)$, or that the average time for $n$ appends is proportional to their write cost (although this is an oversimplification). Then $\dfrac{\FuncTimeDynamicArray(n)}{\FuncTimeGrowthArray(n)} = \dfrac{\FuncWritesDynamicArray(n)}{\FuncWritesGrowthArray(n)}$, where $\FuncWritesDynamicArray$ and $\FuncWritesGrowthArray$ are respectively the write cost functions for dynamic and growth arrays. In this section, I will derive the formula for $\FuncWritesDynamicArray(n)$. Because this section concerns dynamic arrays only, I will denote it $\FuncWrites(n)$.

@@

\begin{lemma}
\label{lem:ToVarUsefulPowerBounds}
	$\dfrac n \VarInitCapacity \FluteLeq \VarGrowthFactor^\VarUseful \FluteLess \dfrac {\VarGrowthFactor{n}} \VarInitCapacity$. That is, $\dfrac n \VarInitCapacity \leq \VarGrowthFactor^\VarUseful < \dfrac {\VarGrowthFactor{n}} \VarInitCapacity$ for sufficiently large $n$.
\end{lemma}

@@

I can build on this inequality to receive asymptotic bounds for $\FuncWrites(n)$:
\begin{align*}
\frac n \VarInitCapacity &\FluteLeq \VarGrowthFactor^\VarUseful \FluteLess \frac {\VarGrowthFactor{n}} \VarInitCapacity\\
\frac n \VarInitCapacity &\FluteLeq \VarGrowthFactor^\VarUseful - 1 \FluteLess \frac {\VarGrowthFactor{n}} \VarInitCapacity &&\text{(Corollary \ref{coro:PlusConstant}, Theorem \ref{thm:InterchangeableInInequality})}\\
\frac n {\VarGrowthFactor - 1} &\FluteLeq \left( \frac{\VarGrowthFactor^\VarUseful - 1}{\VarGrowthFactor - 1} \right) \VarInitCapacity \FluteLess \frac {\VarGrowthFactor{n}} {\VarGrowthFactor - 1} &&\text{(Theorem \ref{thm:BothSidesInequality})}\\
\left( \frac{\VarGrowthFactor}{\VarGrowthFactor - 1} \right) n &\FluteLeq \FuncWrites(n) \FluteLess \left( \frac{2\VarGrowthFactor - 1}{\VarGrowthFactor - 1} \right) n &&\text{(Theorem \ref{thm:BothSidesInequalityAdd})}
\end{align*}
This means that, for large $n$, the number of items written to append $n$ items is roughly between the two bounds shown. For example, suppose $\VarInitCapacity = 4$, $\VarGrowthFactor = 2$, and $32$ items are appended. $\FuncGrow$ is called at the sizes $4$, $8$, and $16$, so $4 + 8 + 16 = 28$ items are copied. Counting the $32$ writes made by $\FuncAppend$, one per item, the write cost is $\FuncWrites(32) = 28 + 32 = 60$. This is approximately equal to the lower bound, $\dfrac{2}{2 - 1} \cdot 32 = 64$.

Now, suppose that one more item is added so that $n = 33$. $\FuncGrow$ is called again, so the number of items copied becomes $4 + 8 + 16 + 32 = 60$. The write cost becomes $\FuncWrites(33) = 33 + 60 = 93$. This is approximately equal to the upper bound, $\dfrac{2 \cdot 2 - 1}{2 - 1} \cdot 33 = 99$.

Intuitively, it makes sense that $\FuncWrites(n)$ should jump from the lower bound to the upper bound when $n = \VarGrowSeq_i + 1$, since at that point $O(n)$ many items are copied from the old to the new buffer. (In the previous example, $\VarGrowSeq_i + 1$ is $17$.) As new items are added, however, both $n$ and $\FuncWrites(n)$ only increase by $1$, and the ratio $\dfrac {\FuncWrites(n)} n$ slowly dwindles toward the lower bound until $n$ reaches $\VarGrowSeq_{i + 1} + 1$.

@@

First, from the definition of $\VarList.\FieldCapacity$, note that a dynamic array's capacity is the length of the buffer it stores its items in. Then a buffer of length $c$ is allocated at some point if and only if $c \in \VarCapacitySeq$. Then the total length of those buffers is
\begin{align*}
\FuncSpace(n) &= \sum_i {\VarCapacitySeq_i}\\
&= \VarInitCapacity + \VarGrowthFactor\VarInitCapacity + \VarGrowthFactor^2\VarInitCapacity + \ldots + \VarGrowthFactor^\VarUseful\VarInitCapacity\\
&= \left( \frac{\VarGrowthFactor^{\VarUseful + 1} - 1}{\VarGrowthFactor - 1} \right) \VarInitCapacity
\end{align*}
Using Lemma \ref{lem:ToVarUsefulPowerBounds} again, I asymptotically bound $\FuncSpace(n)$:
\begin{align*}
\frac n \VarInitCapacity &\FluteLeq \VarGrowthFactor^\VarUseful \FluteLess \frac {\VarGrowthFactor{n}} \VarInitCapacity\\
\frac {\VarGrowthFactor n} \VarInitCapacity &\FluteLeq \VarGrowthFactor^{\VarUseful + 1} \FluteLess \frac {\VarGrowthFactor^2 n} \VarInitCapacity &&\text{(Theorem \ref{thm:BothSidesInequality})}\\
\frac {\VarGrowthFactor n} \VarInitCapacity &\FluteLeq \VarGrowthFactor^{\VarUseful + 1} - 1 \FluteLess \frac {\VarGrowthFactor^2 n} \VarInitCapacity &&\text{(Corollary \ref{coro:PlusConstant}, Theorem \ref{thm:InterchangeableInInequality})}\\
\left( \frac{\VarGrowthFactor}{\VarGrowthFactor - 1} \right) n &\FluteLeq \FuncSpace(n) \FluteLess \left( \frac{\VarGrowthFactor^2}{\VarGrowthFactor - 1} \right) n &&\text{(Theorem \ref{thm:BothSidesInequality})}
\end{align*}
Using the same intuition from the time complexity section, $\FuncSpace(n)$ approaches the lower bound as $n \to \VarGrowSeq_i$, and jumps to the upper bound when $n = \VarGrowSeq_i + 1$. If $100$ items were appended to a dynamic array with $\VarInitCapacity = 11$ and $\VarGrowthFactor = 3$, for example, the wasted space $s(100) - 100$ would be approximately $\dfrac{3^2}{3 - 1} \cdot 100 - 100 = 350$. If $n$ were $99$, however, it would be approximately $\dfrac{3}{3 - 1} \cdot 99 - 99 = 49.5$.

>>>>

I wish to find the ratio $\frac{\FuncTimeDynamicArray(n)}{\FuncTimeGrowthArray(n)}$, where $\FuncTimeDynamicArray(n)$ and $\FuncTimeGrowthArray(n)$ denote the average time needed to append $n$ items for dynamic and growth arrays, respectively. This represents how much faster growth arrays are at appending than dynamic ones. I will approximate this ratio using the following question: Suppose $n$ elements are appended to an empty collection. How many times is an element stored in an array? I will term the answer to this question the \textbf{write cost} of $n$ appends, and denote it $\FuncWrites(n)$.

I will assume that $\FuncTime(n) \propto \FuncWrites(n)$, or that the average time for $n$ appends is proportional to their write cost (although this is an oversimplification). Then $\frac{\FuncTimeDynamicArray(n)}{\FuncTimeGrowthArray(n)} = \frac{\FuncWritesDynamicArray(n)}{\FuncWritesGrowthArray(n)}$, where $\FuncWritesDynamicArray$ and $\FuncWritesGrowthArray$ are respectively the write cost functions for dynamic and growth arrays. In this section, I will derive the formula for $\FuncWritesDynamicArray(n)$. Because this section concerns dynamic arrays only, I will denote it $\FuncWrites(n)$.

@@

\begin{lemma}
\label{lem:ToVarUsefulPowerBounds}
	$\frac n \VarInitCapacity \FluteLeq \VarGrowthFactor^\VarUseful \FluteLess \frac {\VarGrowthFactor{n}} \VarInitCapacity$. That is, $\frac n \VarInitCapacity \leq \VarGrowthFactor^\VarUseful < \frac {\VarGrowthFactor{n}} \VarInitCapacity$ for sufficiently large $n$.
\end{lemma}

@@

I can build on this inequality to receive asymptotic bounds for $\FuncWrites(n)$:
\begin{align*}
\frac n \VarInitCapacity &\FluteLeq \VarGrowthFactor^\VarUseful \FluteLess \frac {\VarGrowthFactor{n}} \VarInitCapacity\\
\frac n \VarInitCapacity &\FluteLeq \VarGrowthFactor^\VarUseful - 1 \FluteLess \frac {\VarGrowthFactor{n}} \VarInitCapacity &&\text{(Corollary \ref{coro:PlusConstant}, Theorem \ref{thm:InterchangeableInInequality})}\\
\frac n {\VarGrowthFactor - 1} &\FluteLeq \left( \frac{\VarGrowthFactor^\VarUseful - 1}{\VarGrowthFactor - 1} \right) \VarInitCapacity \FluteLess \frac {\VarGrowthFactor{n}} {\VarGrowthFactor - 1} &&\text{(Theorem \ref{thm:BothSidesInequality})}\\
\left( \frac{\VarGrowthFactor}{\VarGrowthFactor - 1} \right) n &\FluteLeq \FuncWrites(n) \FluteLess \left( \frac{2\VarGrowthFactor - 1}{\VarGrowthFactor - 1} \right) n &&\text{(Theorem \ref{thm:BothSidesInequalityAdd})}
\end{align*}
This means that, for large $n$, the number of items written to append $n$ items is roughly between the two bounds shown. For example, suppose $\VarInitCapacity = 4$, $\VarGrowthFactor = 2$, and $32$ items are appended. $\FuncGrow$ is called at the sizes $4$, $8$, and $16$, so $4 + 8 + 16 = 28$ items are copied. Counting the $32$ writes made by $\FuncAppend$, one per item, the write cost is $\FuncWrites(32) = 28 + 32 = 60$. This is approximately equal to the lower bound, $\frac{2}{2 - 1} \cdot 32 = 64$.

Now, suppose that one more item is added so that $n = 33$. $\FuncGrow$ is called again, so the number of items copied becomes $4 + 8 + 16 + 32 = 60$. The write cost becomes $\FuncWrites(33) = 33 + 60 = 93$. This is approximately equal to the upper bound, $\frac{2 \cdot 2 - 1}{2 - 1} \cdot 33 = 99$.

Intuitively, it makes sense that $\FuncWrites(n)$ should jump from the lower bound to the upper bound when $n = \VarGrowSeq_i + 1$, since at that point $O(n)$ many items are copied from the old to the new buffer. (In the previous example, $\VarGrowSeq_i + 1$ is $17$.) As new items are added, however, both $n$ and $\FuncWrites(n)$ only increase by $1$, and the ratio $\frac {\FuncWrites(n)} n$ slowly dwindles toward the lower bound until $n$ reaches $\VarGrowSeq_{i + 1} + 1$.

@@

First, from the definition of $\VarList.\FieldCapacity$, note that a dynamic array's capacity is the length of the buffer it stores its items in. Then a buffer of length $c$ is allocated at some point if and only if $c \in \VarCapacitySeq$. Then the total length of those buffers is
\begin{align*}
\FuncSpace(n) &= \sum_i {\VarCapacitySeq_i}\\
&= \VarInitCapacity + \VarGrowthFactor\VarInitCapacity + \VarGrowthFactor^2\VarInitCapacity + \ldots + \VarGrowthFactor^\VarUseful\VarInitCapacity\\
&= \left( \frac{\VarGrowthFactor^{\VarUseful + 1} - 1}{\VarGrowthFactor - 1} \right) \VarInitCapacity
\end{align*}
Using Lemma \ref{lem:ToVarUsefulPowerBounds} again, I asymptotically bound $\FuncSpace(n)$:
\begin{align*}
\frac n \VarInitCapacity &\FluteLeq \VarGrowthFactor^\VarUseful \FluteLess \frac {\VarGrowthFactor{n}} \VarInitCapacity\\
\frac {\VarGrowthFactor n} \VarInitCapacity &\FluteLeq \VarGrowthFactor^{\VarUseful + 1} \FluteLess \frac {\VarGrowthFactor^2 n} \VarInitCapacity &&\text{(Theorem \ref{thm:BothSidesInequality})}\\
\frac {\VarGrowthFactor n} \VarInitCapacity &\FluteLeq \VarGrowthFactor^{\VarUseful + 1} - 1 \FluteLess \frac {\VarGrowthFactor^2 n} \VarInitCapacity &&\text{(Corollary \ref{coro:PlusConstant}, Theorem \ref{thm:InterchangeableInInequality})}\\
\left( \frac{\VarGrowthFactor}{\VarGrowthFactor - 1} \right) n &\FluteLeq \FuncSpace(n) \FluteLess \left( \frac{\VarGrowthFactor^2}{\VarGrowthFactor - 1} \right) n &&\text{(Theorem \ref{thm:BothSidesInequality})}
\end{align*}
Using the same intuition from the time complexity section, $\FuncSpace(n)$ approaches the lower bound as $n \to \VarGrowSeq_i$, and jumps to the upper bound when $n = \VarGrowSeq_i + 1$. If $100$ items were appended to a dynamic array with $\VarInitCapacity = 11$ and $\VarGrowthFactor = 3$, for example, the wasted space $s(100) - 100$ would be approximately $\frac{3^2}{3 - 1} \cdot 100 - 100 = 350$. If $n$ were $99$, however, it would be approximately $\frac{3}{3 - 1} \cdot 99 - 99 = 49.5$.
